---
output:
  html_document:
    toc: true
---

# compare 3-mer mutation rate model and 7-mer mutation rate model

```{r}
library(data.table)
library(parallel)
source("../TADA-A/lib/TADA_Annotation.R")
source("../lib/TADA-A-plus.R")
```

## denovo_db_DeRubeis2014_cases_SNV_remove_recurrent_mutations

```{r, cache=TRUE, cache.extra=TADA_A_adjust_mutation_rate}
chr <- paste("chr", seq(1,22),sep = "")
output_model_1 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)
output_model_2 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)

for(i in 1:10){
# Use all chromosomes except chr1 to train mutation rate model, only use nonTADA_A genes to train mutation model.
  a <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_DeRubeis2014_cases_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment.bed",
                              sample_size = 2270, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_1[i,] <- as.vector(a)

  b <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_DeRubeis2014_cases_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment_ERV_mutrate.bed",
                              sample_size = 2270, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_2[i,] <- as.vector(b)
}
```

### 3-mer model
```{r}
knitr::kable(output_model_1)
```

### 7-mer model
```{r}
knitr::kable(output_model_2)
```

### comparison of loglikelihood
```{r}
boxplot(output_model_1$ll - output_model_2$ll)
```



## denovo_db_Krumm_cases_SNV_remove_recurrent_mutations

```{r, cache=TRUE, cache.extra=TADA_A_adjust_mutation_rate}
chr <- paste("chr", seq(1,22),sep = "")
output_model_1 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)
output_model_2 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)

for(i in 1:10){
# Use all chromosomes except chr1 to train mutation rate model, only use nonTADA_A genes to train mutation model.
  a <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Krumm_cases_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment.bed",
                              sample_size = 2377, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_1[i,] <- as.vector(a)

  b <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Krumm_cases_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment_ERV_mutrate.bed",
                              sample_size = 2377, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_2[i,] <- as.vector(b)
}
```

### 3-mer model
```{r}
knitr::kable(output_model_1)
```

### 7-mer model
```{r}
knitr::kable(output_model_2)
```

### comparison of loglikelihood
```{r}
boxplot(output_model_1$ll - output_model_2$ll)
```


## denovo_db_Krumm_controls_SNV_remove_recurrent_mutations

```{r, cache=TRUE, cache.extra=TADA_A_adjust_mutation_rate}
chr <- paste("chr", seq(1,22),sep = "")
output_model_1 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)
output_model_2 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)

for(i in 1:10){
# Use all chromosomes except chr1 to train mutation rate model, only use nonTADA_A genes to train mutation model.
  a <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Krumm_controls_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment.bed",
                              sample_size = 1786, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_1[i,] <- as.vector(a)

  b <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Krumm_controls_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment_ERV_mutrate.bed",
                              sample_size = 1786, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_2[i,] <- as.vector(b)
}
```

### 3-mer model
```{r}
knitr::kable(output_model_1)
```

### 7-mer model
```{r}
knitr::kable(output_model_2)
```

### comparison of loglikelihood
```{r}
boxplot(output_model_1$ll - output_model_2$ll)
```

## denovo_db_Iossifov_cases_SNV_remove_recurrent_mutations

```{r, cache=TRUE, cache.extra=TADA_A_adjust_mutation_rate}
chr <- paste("chr", seq(1,22),sep = "")
output_model_1 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)
output_model_2 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)

for(i in 1:10){
# Use all chromosomes except chr1 to train mutation rate model, only use nonTADA_A genes to train mutation model.
  a <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Iossifov_cases_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment.bed",
                              sample_size = 2508, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_1[i,] <- as.vector(a)

  b <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Iossifov_cases_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment_ERV_mutrate.bed",
                              sample_size = 2508, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_2[i,] <- as.vector(b)
}
```

### 3-mer model
```{r}
knitr::kable(output_model_1)
```

### 7-mer model
```{r}
knitr::kable(output_model_2)
```

### comparison of loglikelihood
```{r}
boxplot(output_model_1$ll - output_model_2$ll)
```


## denovo_db_Iossifov_controls_SNV_remove_recurrent_mutations

```{r, cache=TRUE, cache.extra=TADA_A_adjust_mutation_rate}
chr <- paste("chr", seq(1,22),sep = "")
output_model_1 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)
output_model_2 <- data.frame(ll = NA, expected_count = NA, observed_count = NA)

for(i in 1:10){
# Use all chromosomes except chr1 to train mutation rate model, only use nonTADA_A genes to train mutation model.
  a <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Iossifov_controls_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment.bed",
                              sample_size = 1911, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_1[i,] <- as.vector(a)

  b <- TADA_A_compare_mutrate_model(mut_file = "../data/denovo_db_Iossifov_controls_SNV_remove_recurrent_mutations.bed",
                              window_file = "../data/Example_windows_coding_for_mutation_rate_adjustment_ERV_mutrate.bed",
                              sample_size = 1911, 
                              scale_features = c("GC_content", "div_score"),
                              validation_chr = chr[i],
                              mutrate_mode = "special",
                              genes = "../data/Example_gene_prior_st_0.05.txt")
  output_model_2[i,] <- as.vector(b)
}
```

### 3-mer model
```{r}
knitr::kable(output_model_1)
```

### 7-mer model
```{r}
knitr::kable(output_model_2)
```

### comparison of loglikelihood
```{r}
boxplot(output_model_1$ll - output_model_2$ll)
```
